# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r}
activity <- read.csv("activity.csv")
activity_complete <- activity[complete.cases(activity),] # strip out NAs
```


## What is mean total number of steps taken per day?

```{r}
total_steps <- aggregate(activity_complete$steps, list(activity_complete$date), sum)
hist(total_steps$x, main="Total Steps per Day Oct + Nov 2012",xlab="Steps")
mean_ts<-mean(total_steps$x)
med_ts<-median(total_steps$x)
```

The mean of daily total steps in sample period is `r mean_ts`.

The median of daily total steps in sample period is `r med_ts`.

## What is the average daily activity pattern?

```{r}
library (ggplot2)
mean_steps <- aggregate(activity_complete$steps, list(activity_complete$interval), mean)

d<-ggplot(mean_steps,aes(Group.1,x))
d<-d+geom_line()+xlab("Interval")+ylab("Mean Steps")+ggtitle("Mean Steps October + November 2012")
d

# determine max number of steps during any period and the period inverval
max_mean_steps<-0
max_interval<-0
for (i in 1:nrow(mean_steps))
  if(mean_steps[i,2]>max_mean_steps) {
    max_mean_steps<-mean_steps[i,2]
    max_interval<-mean_steps[i,1]
}

```

Maximum number of steps was `r max_mean_steps` at time interval `r max_interval`.


## Imputing missing values

```{r}
missing <- nrow(activity[!complete.cases(activity),])
```

There are `r missing` rows containing missing data

Filling in missing data
Determine Means for each time interval
```{r}
activity_imp <- activity  # create a new data frame to hold activity data with imputations
# using existing data, create median number of steps per time interval
intervals = unique(activity$interval)
nintervals = length(intervals)
meanstep = numeric(nintervals)
for(i in 1:nintervals)
  meanstep[i]=median(activity$steps[activity$interval==intervals[i]],na.rm=TRUE)
# use the median data collected above to fill the NAs
for(i in 1:dim(activity)[1])
  if (is.na(activity[i,1]))
    activity_imp[i,1]=meanstep[which(intervals==activity[i,3])]
# determine if missing data remains
missing <- nrow(activity_imp[!complete.cases(activity_imp),])

# create new histogram, mean and median statistics from imputed data
total_steps <- aggregate(activity_imp$steps, list(activity_imp$date), sum)
hist(total_steps$x, main="Total Steps per Day Oct + Nov 2012 (Imputed Data)",xlab="Steps")
mean_ts<-mean(total_steps$x)
med_ts<-median(total_steps$x)

# regenerate profile of daily activity (won't be too different if we use mean or median to fill data)
mean_steps <- aggregate(activity_imp$steps, list(activity_imp$interval), mean)
d<-ggplot(mean_steps,aes(Group.1,x))
d<-d+geom_line()+xlab("Interval")+ylab("Mean Steps")+ggtitle("Mean Steps October + November 2012 (Imputed Data Set)")
d
```

There are `r missing` rows containing missing data in the imputed data set.

The mean of daily total steps in sample period imputing missing data is `r mean_ts`.

The median of daily total steps in sample period imputing missing data is `r med_ts`.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
# categorize down weekend and weekday observations
activity_complete$weekday = weekdays(as.Date(activity_complete$date))
activity_complete["day_type"] <- NA
activity_complete$day_type[!(activity_complete$weekday == "Saturday
" | activity_complete$weekday == "Sunday")] <- "Weekday"
activity_complete$day_type[activity_complete$weekday == "Saturday" | activity_complete$weekday == "Sunday"] <- "Weekend"
# build weekend data set
mean_steps_we <- aggregate(activity_complete$steps[activity_complete$day_type=="Weekend"], list(activity_complete$interval[activity_complete$day_type=="Weekend"]), mean)
# build weekday data set 
mean_steps_wd <- aggregate(activity_complete$steps[activity_complete$day_type=="Weekday"], list(activity_complete$interval[activity_complete$day_type=="Weekday"]), mean)

# plot both datasets together

d<-ggplot(mean_steps_we,aes(Group.1,x))
d<-d+geom_line()+xlab("Interval")+ylab("Mean Steps")+ggtitle("Weekend")
d2<-ggplot(mean_steps_wd,aes(Group.1,x))
d2<-d+geom_line()+xlab("Interval")+ylab("Mean Steps")+ggtitle("Weekday")
library(grid)
pushViewport(viewport(layout = grid.layout(2,1)))
print(d, vp = viewport(layout.pos.row = 1, layout.pos.col = 1))
print(d2, vp = viewport(layout.pos.row = 2, layout.pos.col = 1))
```


